SI

### 2021-02-08 ms/si.Rmd
Add NaCl sporo chamber plot

plot_sporo_chamber_time <- function (data, time_limits) {
    ggplot(data, aes(x = species, ymin = pre, ymax = post, y = post, group = species_duration, linetype = dry_time)) +
    geom_linerange(position = position_dodge(0.5)) +
    geom_point(position = position_dodge(0.5)) +
    scale_y_datetime(limits = time_limits) +
    scale_linetype_manual(values = c("15" = "solid", "2" = "solid")) +
    coord_flip() +
    labs(x = "Species") +
    theme(
      axis.title.x = element_blank(),
      legend.position = "none"
    )
}

@@

```{r sporophyte-dt-2013-prep}
# Extract pre- and post-DT treatment times for sporophytes
sporo_times <-
  filmy_dt %>%
  filter(generation == "sporophyte") %>%
  select(species:dataset, contains("time_")) %>%
  pivot_longer(names_to = "condition", values_to = "date_time", contains("time_")) %>%
  filter(!is.na(date_time)) %>%
  mutate(condition = str_remove_all(condition, "time_") %>%
           # Consider first recovery measurement (at 30 min) to be "post-DT treatment"
           str_replace_all("30min", "post")) %>%
  filter(condition %in% c("pre", "post")) %>%
  add_count(species, salt, dry_time, individual, dataset) %>%
  filter(n > 1) %>%
  select(-n) %>%
  pivot_wider(names_from = "condition", values_from = "date_time")
```

```{r sporophyte-dt-2013-LiCl, fig.height = 4.5, fig.width = 7, fig.cap = ""}
# Get mean times of each cluster of DT test samples for the selected drying salt
sporo_clusters_2013 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2013")), "LiCl", 3)
sporo_clusters_2014 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2014")), "LiCl", 1)

sporo_clusters <- bind_rows(sporo_clusters_2013, sporo_clusters_2014)

# Calculate time range for the clustered data
sporo_time_range_2013 <- c(min = min(sporo_clusters_2013$pre), max = max(sporo_clusters_2013$post))
sporo_time_range_2014 <- c(min = min(sporo_clusters_2014$pre), max = max(sporo_clusters_2014$post))

# Subset DT chamber data
licl_chamber_data_2013 <-
  filter(filmy_dt_chamber, 
    date_time >= sporo_time_range_2013[["min"]], 
    date_time <= sporo_time_range_2013[["max"]], 
    salt == "LiCl")
licl_chamber_data_2014 <- 
  filter(filmy_dt_chamber, 
    date_time >= sporo_time_range_2014[["min"]], 
    date_time <= sporo_time_range_2014[["max"]], 
    salt == "LiCl") %>%
  # There are two sets of times: 2014-06-17 to 2014-07-02, and 2014-07-06 to 2014-07-08
  # Add a group for these so ggplot doesn't connect lines betweeen them
  mutate(
    day = yday(date_time),
    group = ifelse(day < 185, "a", "b"))

@@

# RH plot
rh_plot_2013 <- plot_chamber_rh(licl_chamber_data_2013, sporo_time_range_2013) +
  scale_y_continuous(limits = licl_rh_range, label = label_percent(accuracy = 1))

rh_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = 0.01*rh, group = group)) +
  geom_line() +
  scale_x_datetime(limits = sporo_time_range_2014) +
  blank_x_theme() +
  blank_y_theme() +
  theme(legend.position = "none")

# Temp plot
temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, sporo_time_range_2013) +
  scale_y_continuous(limits = licl_temp_range, label = label_number(suffix = "\u00b0C", accuracy = 1))

temp_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = temp, group = group)) +
  geom_line() +
  scale_x_datetime(limits = sporo_time_range_2014) +
  blank_x_theme() +
  blank_y_theme() +
  theme(legend.position = "none")

# Dry times plot
time_plot_2013 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2013) +
  labs(y = "2013") +
  theme(axis.title.x = element_text()) 
time_plot_2014 <- plot_sporo_chamber_time(sporo_clusters, sporo_time_range_2014)  +
  labs(y = "2014")  +
  theme(axis.title.x = element_text()) +
  blank_y_theme()

# Combine plots
temp_plot_2013 + temp_plot_2014 + 
  rh_plot_2013 + rh_plot_2014 + 
  time_plot_2013 + time_plot_2014 + 
  plot_layout(ncol = 2)

# ggsave(here::here("results/S2_fig.png"))
```

**`r s_figure("chamber-s")`**. Conditions inside desiccation chamber and duration of desiccation treatment for sporophytes. LiCl was used to maintain constant humidity at \circa{ }18%.

>>>>

plot_sporo_chamber_time <- function (data, time_limits) {
    ggplot(data, aes(x = species, ymin = pre, ymax = post, y = post, group = species_duration, linetype = dry_time)) +
    geom_linerange(position = position_dodge(0.5)) +
    geom_point(position = position_dodge(0.5)) +
    scale_y_datetime(limits = time_limits) +
    scale_x_discrete(labels = str_abbrev_sp_pretty) +
    scale_linetype_manual(values = c("15" = "solid", "2" = "solid")) +
    coord_flip() +
    labs(x = "Species") +
    theme(
      axis.title.x = element_blank(),
      axis.text.y = element_markdown(),
      legend.position = "none"
    )
}

@@

```{r sporophyte-dt-prep}
# Extract pre- and post-DT treatment times for sporophytes
sporo_times <-
  filmy_dt %>%
  filter(generation == "sporophyte") %>%
  select(species:dataset, contains("time_")) %>%
  pivot_longer(names_to = "condition", values_to = "date_time", contains("time_")) %>%
  filter(!is.na(date_time)) %>%
  mutate(condition = str_remove_all(condition, "time_") %>%
           # Consider first recovery measurement (at 30 min) to be "post-DT treatment"
           str_replace_all("30min", "post")) %>%
  filter(condition %in% c("pre", "post")) %>%
  add_count(species, salt, dry_time, individual, dataset) %>%
  filter(n > 1) %>%  # only keep those with both pre- and post- times
  select(-n) %>%
  pivot_wider(names_from = "condition", values_from = "date_time")
```

```{r sporophyte-dt-2013-LiCl, fig.height = 4.5, fig.width = 7, fig.cap = ""}
# Get mean times of each cluster of DT test samples for the selected drying salt
licl_clusters_2013 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2013")), "LiCl", 3)
licl_clusters_2014 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2014")), "LiCl", 1)

licl_clusters <- bind_rows(licl_clusters_2013, licl_clusters_2014)

# Calculate time range for the clustered data
licl_time_range_2013 <- c(min = min(licl_clusters_2013$pre), max = max(licl_clusters_2013$post))
licl_time_range_2014 <- c(min = min(licl_clusters_2014$pre), max = max(licl_clusters_2014$post))

# Subset DT chamber data
licl_chamber_data_2013 <-
  filter(filmy_dt_chamber, 
    date_time >= licl_time_range_2013[["min"]], 
    date_time <= licl_time_range_2013[["max"]], 
    salt == "LiCl")
licl_chamber_data_2014 <- 
  filter(filmy_dt_chamber, 
    date_time >= licl_time_range_2014[["min"]], 
    date_time <= licl_time_range_2014[["max"]], 
    salt == "LiCl") %>%
  # There are two sets of times: 2014-06-17 to 2014-07-02, and 2014-07-06 to 2014-07-08
  # Add a group for these so ggplot doesn't connect lines betweeen them
  mutate(
    day = yday(date_time),
    group = ifelse(day < 185, "a", "b"))

@@

# RH plot
licl_rh_plot_2013 <- plot_chamber_rh(licl_chamber_data_2013, licl_time_range_2013) +
  scale_y_continuous(limits = licl_rh_range, label = label_percent(accuracy = 1))

licl_rh_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = 0.01*rh, group = group)) +
  geom_line() +
  scale_x_datetime(limits = licl_time_range_2014) +
  blank_x_theme() +
  blank_y_theme() +
  theme(legend.position = "none")

# Temp plot
licl_temp_plot_2013 <- plot_chamber_temp(licl_chamber_data_2013, licl_time_range_2013) +
  scale_y_continuous(limits = licl_temp_range, label = label_number(suffix = "\u00b0C", accuracy = 1))

licl_temp_plot_2014 <- ggplot(licl_chamber_data_2014, aes(x = date_time, y = temp, group = group)) +
  geom_line() +
  scale_x_datetime(limits = licl_time_range_2014) +
  blank_x_theme() +
  blank_y_theme() +
  theme(legend.position = "none")

# Dry times plot
licl_time_plot_2013 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2013) +
  labs(y = "2013") +
  theme(axis.title.x = element_text()) 
licl_time_plot_2014 <- plot_sporo_chamber_time(licl_clusters, licl_time_range_2014)  +
  labs(y = "2014")  +
  theme(axis.title.x = element_text()) +
  blank_y_theme()

# Combine plots
licl_temp_plot_2013 + licl_temp_plot_2014 + 
  licl_rh_plot_2013 + licl_rh_plot_2014 + 
  licl_time_plot_2013 + licl_time_plot_2014 + 
  plot_layout(ncol = 2) +
  plot_annotation(title = "a")
```

```{r sporophyte-dt-2013-NaCl, fig.height = 7, fig.width = 7, fig.cap = ""}
# Subset desiccation duration data
# H_polyanthos shows n=1, but this is only because pre/post were available for only one sample
# (actually had 8 samples total)
nacl_clusters_2013 <- 
  sporo_times %>%
  filter(str_detect(dataset, "2013")) %>%
  cluster_sporo_times("NaCl", 4) %>%
  mutate(
    serial_no = case_when(
      species %in% c("Didymoglossum_tahitense", "Crepidomanes_kurzii", "Callistopteris_apiifolia") & dry_time == "2" ~ "missing",
      species %in% c("Hymenophyllum_multifidum", "Hymenophyllum_digitatum") ~ "10140667",
      species %in% c("Hymenophyllum_polyanthos") ~ "10152774",
      species %in% c("Abrodictyum_dentatum") & dry_time == "2" ~ "missing",
      TRUE ~ "missing"
    ),
    serial_no = as.factor(serial_no)
  )

nacl_clusters_2014 <-
  sporo_times %>%
  filter(str_detect(dataset, "2014")) %>%
  cluster_sporo_times("NaCl", 1) %>%
  mutate(
    serial_no = case_when(
      species %in% c("Vandenboschia_maxima") & dry_time == "2" ~ "10140661",
      TRUE ~ "10152773"
    ),
    serial_no = as.factor(serial_no)
  )

# make levels of `serial_no` and `species` the same across datasets
nacl_all_sn <- filmy_dt_chamber %>% 
  filter(salt == "NaCl") %>% 
  pull(serial_no) %>% 
  unique %>% 
  sort %>%
  c("missing")

nacl_all_species <- sporo_times %>% 
  filter(salt == "NaCl") %>% 
  pull(species) %>% 
  unique %>% 
  sort

nacl_clusters_2013 <- mutate(
  nacl_clusters_2013, 
  # sort SN in numeric order
  serial_no = fct_expand(serial_no, nacl_all_sn) %>% fct_inseq,
  # sort species in reverse-alphabetical order (so 'A' is on top in plot)
  species = fct_expand(species, nacl_all_species) %>% fct_relevel(function (x) sort(x, decreasing = TRUE))
  )

nacl_clusters_2014 <- mutate(
  nacl_clusters_2014, 
  serial_no = fct_expand(serial_no, nacl_all_sn) %>% fct_inseq,
  species = fct_expand(species, nacl_all_species) %>% fct_relevel(function (x) sort(x, decreasing = TRUE))
  )

# Calculate time range for the clustered data
# to standardize x-axis
nacl_time_range_2013 <- c(min = min(nacl_clusters_2013$pre), max = max(nacl_clusters_2013$post))
nacl_time_range_2014 <- c(min = min(nacl_clusters_2014$pre), max = max(nacl_clusters_2014$post))

# Subset DT chamber data
nacl_chamber_data_2012 <- 
filmy_dt_chamber %>% 
  filter(salt == "NaCl", date_time > "2012-07-31 18:10:00 UTC", date_time < "2013-01-01 00:10:00 UTC")

nacl_time_range_2012 <- c(min = min(nacl_chamber_data_2012$date_time), max = max(nacl_chamber_data_2012$date_time))

nacl_chamber_data_2013 <-
  filter(filmy_dt_chamber, 
         date_time >= nacl_time_range_2013[["min"]], 
         date_time <= nacl_time_range_2013[["max"]], 
         salt == "NaCl")

nacl_chamber_data_2014 <- 
  filter(filmy_dt_chamber, 
         date_time >= nacl_time_range_2014[["min"]], 
         date_time <= nacl_time_range_2014[["max"]], 
         salt == "NaCl") %>%
  mutate(
    # There are two sets of times: 2014-06-17 to 2014-07-02, and 2014-07-06 to 2014-07-08
    # Add a group for these so ggplot doesn't connect lines betweeen them
    day = yday(date_time),
    group = ifelse(day < 185, "a", "b")
  )

# Get overall range of temperature and RH for NaCl
# to standardize y-axis
nacl_temp_range <- filter(filmy_dt_chamber, salt == "NaCl") %>%
  summarize(min = min(temp), max = max(temp)) %>%
  pivot_longer(everything()) %>%
  arrange(value) %>%
  pull(value)

nacl_rh_range <- filter(filmy_dt_chamber, salt == "NaCl") %>%
  mutate(rh = rh * 0.01) %>%
  summarize(min = min(rh), max = max(rh)) %>%
  pivot_longer(everything()) %>%
  arrange(value) %>%
  pull(value)

# Set up color palette for serial numbers
serial_no_cols <- c(
  brewer_pal(palette = "Dark2")(5),
  "grey20") %>%
  set_names(
    bind_rows(nacl_chamber_data_2013, nacl_chamber_data_2014) %>%
      filter(serial_no != "missing") %>%
      pull(serial_no) %>%
      unique %>%
      sort %>%
      c("missing")
  )

# Time plots
nacl_time_plot_2013 <- 
  ggplot(
    nacl_clusters_2013, 
    aes(
      x = species, ymin = pre, ymax = post, y = post, 
      group = species_duration, color = serial_no)
  ) +
  geom_linerange(position = position_dodge(0.5)) +
  geom_point(position = position_dodge(0.5)) +
  scale_y_datetime(limits = nacl_time_range_2013) +
  # Print all 'species' on the y-axis using `drop = false`
  scale_x_discrete(labels = str_abbrev_sp_pretty, drop = FALSE) +
  # Print all levels of 'serial no' in the legend using `drop = false`
  scale_color_manual(values = serial_no_cols, drop = FALSE) +
  coord_flip() +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank())

nacl_time_plot_2014 <- 
  ggplot(
    nacl_clusters_2014, 
    aes(x = species, ymin = pre, ymax = post, y = post, 
        group = species_duration, color = serial_no)
  ) +
  geom_linerange(position = position_dodge(0.5)) +
  geom_point(position = position_dodge(0.5)) +
  scale_y_datetime(limits = nacl_time_range_2014) +
  scale_x_discrete(labels = str_abbrev_sp_pretty, drop = FALSE) +
  scale_color_manual(values = serial_no_cols) +
  coord_flip() +
  labs(x = "Species") +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_markdown()) +
  guides(color = "none")

# RH plots
nacl_rh_plot_2012 <- 
  ggplot(
    nacl_chamber_data_2012, 
    aes(x = date_time, y = 0.01*rh)
  ) +
  geom_line(size = 0.25) +
  scale_x_datetime(limits = nacl_time_range_2012) +
  scale_y_continuous(
    label = scales::label_percent(accuracy = 1), 
    limits = nacl_rh_range) +
  labs(
    y = "Relative humidity",
    color = "Data logger\nserial no.") +
  guides(color = "none") +
  theme(axis.title.x = element_blank()) +
  blank_y_theme()

nacl_rh_plot_2013 <- 
  ggplot(
    nacl_chamber_data_2013, 
    aes(x = date_time, y = 0.01*rh, color = serial_no)
  ) +
  scale_color_manual(values = serial_no_cols) +
  geom_line(size = 0.25) +
  scale_x_datetime(limits = nacl_time_range_2013) +
  scale_y_continuous(
    label = scales::label_percent(accuracy = 1), 
    limits = nacl_rh_range) +
  labs(
    y = "Relative humidity",
    color = "Data logger\nserial no.") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  blank_y_theme() +
  guides(color = "none")

nacl_rh_plot_2014 <- 
  ggplot(
    nacl_chamber_data_2014, 
    aes(x = date_time, y = 0.01*rh, color = serial_no, group = interaction(group, serial_no))
  ) +
  scale_color_manual(values = serial_no_cols) +
  geom_line(size = 0.25) +
  scale_x_datetime(limits = nacl_time_range_2014) +
  scale_y_continuous(
    label = scales::label_percent(accuracy = 1), 
    limits = nacl_rh_range) +
  labs(
    y = "Relative humidity",
    color = "Data logger\nserial no.") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  guides(color = "none")

# Temp plots
nacl_temp_plot_2012 <- 
  ggplot(
    nacl_chamber_data_2012, 
    aes(x = date_time, y = temp)
  ) +
  geom_line(size = 0.25) +
  scale_x_datetime(limits = nacl_time_range_2012) +
  scale_y_continuous(
    label = label_number(suffix = "\u00b0C", accuracy = 1), 
    limits = nacl_temp_range) +
  labs(
    y = "Temperature",
    color = "Data logger\nserial no.",
    title = "2012") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  blank_y_theme() +
  guides(color = "none")

nacl_temp_plot_2013 <- 
  ggplot(
    nacl_chamber_data_2013, 
    aes(x = date_time, y = temp, color = serial_no)
  ) +
  geom_line(size = 0.25) +
  scale_color_manual(values = serial_no_cols) +
  scale_x_datetime(limits = nacl_time_range_2013) +
  scale_y_continuous(
    label = label_number(suffix = "\u00b0C", accuracy = 1), 
    limits = nacl_temp_range) +
  labs(
    y = "Temperature",
    color = "Data logger\nserial no.",
    title = "2013") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  blank_y_theme() +
  guides(color = "none")

nacl_temp_plot_2014 <- 
  ggplot(
    nacl_chamber_data_2014, 
    aes(x = date_time, y = temp, color = serial_no, group = interaction(group, serial_no))
  ) +
  scale_color_manual(values = serial_no_cols) +
  geom_line(size = 0.25) +
  scale_x_datetime(limits = nacl_time_range_2014) +
  scale_y_continuous(
    label = label_number(suffix = "\u00b0C", accuracy = 1), 
    limits = nacl_temp_range) +
  labs(
    y = "Temperature",
    color = "Data logger\nserial no.",
    title = "2014") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  guides(color = "none")

nacl_temp_plot_2014 + nacl_temp_plot_2013 + nacl_temp_plot_2012 +
  nacl_rh_plot_2014 + nacl_rh_plot_2013 + nacl_rh_plot_2012 +
  nacl_time_plot_2014 + nacl_time_plot_2013 + plot_spacer() +
  plot_layout(ncol = 3, nrow = 3, guides = "collect") +
  plot_annotation(title = "b") &
  theme(legend.position = "bottom")
```

**`r s_figure("chamber-s")`**. Temperature, relative humidity (RH), and duration of desiccation treatment for sporophytes. Relative humidity was controlled by placing drying salts inside desiccation chambers: **a**, LiCl (-282 MPa, 18 % RH); **b**, Mg(NO~3~)~2~ (-86 MPa, 58 % RH); **c**, NaCl (-38 MPa, 80 % RH; **c**. Temperature and RH measured with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) in 2013 and 2014 and **--FIXME: add datalogger model--** dataloggers in 2012. For bottom-most panels (duration of desiccation treatment), line indicates duration and point indicates removal of sample from chamber. For **b** and **c**, multiple chambers were set up and tracked with different data loggers; serial number of data logger indicated by color. "missing" indicates data were not collected on the correspondence between samples and data loggers (chambers).
